---
title: Coursera - Reproducible Research
subtitle: Assignment 1
author: Ron Ammar
output: 
  html_document:
    keep_md: true
    toc: true
    toc_depth: 4
---

```{r global_options, include=FALSE}
  # use include=FALSE to have the chunk evaluated, but neither the code nor its output displayed.
  knitr::opts_chunk$set(echo=TRUE, message=FALSE, fig.align="center",
                        fig.width=12, fig.height=8)
```

```{r}
  # Clear the current session, to avoid errors from persisting data structures
  rm(list=ls())

  # Free up memory by forcing garbage collection
  invisible(gc())

  # Pretty printing in knitr
  library(printr)

  # Manually set the seed to an arbitrary number for consistency in reports
  set.seed(1234)

  # Do not convert character vectors to factors unless explicitly indicated
  options(stringsAsFactors=FALSE)

  startTime <- Sys.time()
```

## Background

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Purpose

The purpose of this analysis is to:

1. Complete the Reproducible Research course project #1.
2. Figure out how active this anonymous dude is.

## Analysis

**Analysis Date:** `r format(Sys.time(), "%k:%M:%S, %A, %B %d, %Y")` 

```{r}
  library(dplyr)
  library(ggplot2)
  
  activity <- read.csv("activity.csv")
  activity$date <- as.Date(activity$date)  # coerce to Date
  
  activityWithMissing <- activity
  # Remove dates with missing step data
  activity <- filter(activity, !is.na(steps))
```

### What is mean total number of steps taken per day?

```{r}
  ggplot(activity, aes(x=date, y=steps)) +
    geom_bar(stat="identity") + 
    theme_bw() + 
    theme(axis.text.x=element_text(angle=90)) +
    labs(y="steps/day")
  
  averages <- activity %>%
    group_by(date) %>%
    summarise(mean_steps=mean(steps),
              median_steps=median(steps))
  averages
```

It would appear that for most of the day, according to each of the measured intervals, this individual is not taking any steps, which is why the median steps is 0 for each day.

For example, if we just look at a few individual days:

```{r}
  threeDays <- activity %>%
    filter(date %in% sample(averages$date, 3)) %>%
    arrange(date)
  
  ggplot(threeDays, aes(x=interval, y=steps, col=date)) +
    facet_grid(date ~ .) +
    geom_line(size=1) +
    theme_bw()
```

This is explained further in the next section.

### What is the average daily activity pattern?

```{r}
  averagePattern <- activity %>%
    group_by(interval) %>%
    summarise(mean_steps=mean(steps))
  
  ggplot(averagePattern, aes(x=interval, y=mean_steps)) +
    geom_line(size=1) +
    theme_bw() +
    labs(y="mean(steps)/interval, summarised across all days")
```

This individual does most of their activity at 5 minute interval #`r averagePattern$interval[which(averagePattern$mean_steps == max(averagePattern$mean_steps))]`.

### Imputing missing values

There are `r nrow(activityWithMissing) - nrow(activity)` missing values in the dataset.

We're going to impute the values that are missing using the suggested method of filling missing values with the mean across all days. However, before proceeding with this method, I need to ensure that there is no interval that is missing for *all* dates. If there is, we'd have to switch to a daily mean strategy.

```{r}
  # If any NaN values remain after averaging, that means a single interval
  # contained all NAs.
  averageWithMissingPattern <- activityWithMissing %>%
    group_by(interval) %>%
    summarise(mean_steps=mean(steps, na.rm=TRUE))
  
  allNA <- any(is.nan(averageWithMissingPattern$mean_steps))
```

Since `allNA` equals ``r allNA``, we can use this mean activity strategy for imputation.

```{r}
  imputedActivty <- activityWithMissing %>%
    filter(is.na(steps)) %>%
    left_join(averageWithMissingPattern, by="interval") %>%
    select(-steps) %>%
    dplyr::rename(steps=mean_steps)
  
  # Add activity data and rearrange
  imputedActivty <- activity %>%
    bind_rows(imputedActivty) %>%
    arrange(date, interval) %>%
    mutate(type="imputed")
```

Now, let's see how the imputed activity differs from the original activity

```{r}
  comparisonActivity <- activity %>%
    arrange(date, interval) %>%  # ensure same order
    mutate(type="original") %>%
    bind_rows(imputedActivty)
  
  ggplot(comparisonActivity, aes(x=date, y=steps)) +
    facet_grid(type ~ .) +
    geom_bar(stat="identity") + 
    theme_bw() + 
    theme(axis.text.x=element_text(angle=90)) +
    labs(y="steps/day")
  
  averagesOriginal <- comparisonActivity %>%
    filter(type == "original") %>%
    group_by(date) %>%
    summarise(mean_steps=mean(steps),
              median_steps=median(steps))
  
  averagesImputed <- comparisonActivity %>%
    filter(type == "imputed") %>%
    group_by(date) %>%
    summarise(imputed_mean_steps=mean(steps),
              imputed_median_steps=median(steps))
  
  averages <- left_join(averagesImputed, averagesOriginal, by="date")
  averages
```

By imputing the activity data, it looks like we can generate rough estimates for activity on days with no activity at all. These numbers appear to be in line with the general activity profile, as plotted in the histograms.

### Are there differences in activity patterns between weekdays and weekends?




------

## System Information

***Time required to process this report:*** *`r format(Sys.time() - startTime)`*

***R session information:***

```{r, echo_session_info}
  sessionInfo()
```